---
title: "chicago"
output:
  html_document: default
  word_document: default
date: "2022-12-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1.loading packages
```{r}
library(tidyverse)
library(tmap)
library(spdep)
library(plotly)
library(rgdal)
library(broom)
library(mapview)
library(crosstalk)
library(sf)
library(sp)
library(spdep)
library(car)
library(fs)
library(janitor)
library(dplyr)
```


# 2. read data
```{r}
# read shapefile
chicago_shape<-st_read(here::here("shape","geo_export_dffff05e-477e-4cf2-b94c-8a3b293c84f4.shp"))

#potential factors for crimes
socioeconomic_factors <- read_csv(here::here("Census_Data_-_Selected_socioeconomic_indicators_in_Chicago__2008___2012.csv"), 
                         na = c(" ", "NA", "n/a"), 
                         col_names = TRUE, 
                         locale = locale(encoding = 'Latin1'))%>%
  clean_names()
#crime data 
chicago_crime <- read_csv(here::here("Crimes2010.csv"), 
                         na = c(" ", "NA", "n/a"), 
                         col_names = TRUE, 
                         locale = locale(encoding = 'Latin1'))%>%
  clean_names()
```

```{r}
qtm(chicago_shape)
```
#3. data processing 
```{r}
socioeconomic_factors<-socioeconomic_factors%>%
  dplyr::filter(., !is.na(community_area_number))
# remove na
chicago_crime<-chicago_crime%>%
  dplyr::filter(., !is.na(longitude))%>%
  dplyr::filter(., !is.na(latitude))
```
```{r}
# check all of the columns have been read in correctly 
Datatypelist1 <- chicago_crime %>% 
  summarise_all(class) %>%
  pivot_longer(everything(), 
               names_to="All_variables", 
               values_to="Variable_class")
Datatypelist1
```

```{r}
#check all of the columns have been read in correctly 
Datatypelist2 <- socioeconomic_factors %>% 
  summarise_all(class) %>%
  pivot_longer(everything(), 
               names_to="All_variables", 
               values_to="Variable_class")

Datatypelist2
```
```{r}
# select burglary data
chicago_crime_reduce<-chicago_crime%>%
#  filter(str_detect(date,"^01"))%>%
  filter(primary_type=="BURGLARY")%>%
  filter(arrest=="TRUE")
```

```{r}
# correct crs
chicago_crime_sf<-st_as_sf(chicago_crime_reduce,
                          coords = c("longitude","latitude"),
                          crs = 4326)
# remove duplicates
chicago_crime_sf<-distinct(chicago_crime_sf)
```

```{r}
# plot crime points on the map 
tmap_mode("plot")
tm_shape(chicago_shape)+
  tm_polygons(col=NA,alpha = 1)+
tm_shape(chicago_crime_sf)+
  tm_dots(col="purple")
```
# make sure all points within Chicago boundary
```{r}
chicago_shape<-chicago_shape%>%
  st_transform(.,26971)
```

```{r}
chicago_crime_sf<-chicago_crime_sf%>%
  st_transform(.,26971)
```

```{r}
chicago_crime_sf<-chicago_crime_sf[chicago_shape,]
```

# calculate density of crime points
```{r}
library(sf)
chicago_points_sf_joined <- chicago_shape%>%
  mutate(n = lengths(st_intersects(., chicago_crime_sf)))%>%
  janitor::clean_names()%>%
  #calculate area
  mutate(area=st_area(.))%>%
  #then density of the points per ward
  mutate(density=n/area)
```


```{r}
# plot density map 
tm_shape(chicago_points_sf_joined) +
    tm_polygons("density",
        style="jenks",
        palette="PuOr",
        midpoint=NA,
        popup.vars=c("wardname", "density"),
        title="Crime Density N/m2")+
  tm_layout(legend.position = c(0.02,0.08))+
  tm_scale_bar(position = c(0.02, 0.01), size = 0.5,text.size = .75)+
  tm_compass(position = c(0.85, 0.8))
```
# point pattern analysis (kernal denisty)
```{r}
library(spatstat)
```

```{r}
#now set a window as the borough boundary
window <- as.owin(chicago_shape)
plot(window)
```
```{r}
#create a sp object
chicago_sub<- chicago_crime_sf %>%
  as(., 'Spatial')
#create a ppp object
chicago_sub.ppp <- ppp(x=chicago_sub@coords[,1],
                          y=chicago_sub@coords[,2],
                          window=window)
```

```{r}
chicago_sub.ppp %>%
  density(., sigma=1000) %>%
  plot()
```

# calculate weight matrix 
```{r}
library(spdep)
```

```{r}
coordsW <- chicago_points_sf_joined%>%
  st_centroid()%>%
  st_geometry()
  
plot(coordsW,axes=TRUE)
#create a neighbours list
LWard_nb <- chicago_points_sf_joined %>%
  poly2nb(., queen = F)

```
```{r}
summary(LWard_nb)
```
```{r}
#plot them
plot(LWard_nb, st_geometry(coordsW), col="red")
#add a map underneath
plot(chicago_points_sf_joined$geometry, add=T)
```

```{r}
Lward.lw <- LWard_nb %>%
  nb2listw(., style="W")
```

# Global Moran's I for density of crimes 
```{r}
I_LWard_Global_Density <- chicago_points_sf_joined %>%
  pull(density) %>%
  as.vector()%>%
  moran.test(., Lward.lw)

I_LWard_Global_Density
```
# Local Moran's I for density of crimes 
```{r}
I_LWard_Local_Density <- chicago_points_sf_joined %>%
  pull(density) %>%
  as.vector()%>%
  localmoran(., Lward.lw)%>%
  as_tibble()
slice_head(I_LWard_Local_Density, n=5)
```

```{r}
chicago_points_sf_joined <- chicago_points_sf_joined %>%
  mutate(density_I =as.numeric(I_LWard_Local_Density$Ii))%>%
  mutate(density_Iz =as.numeric(I_LWard_Local_Density$Z.Ii))
```

```{r}
breaks1<-c(-1000,-2.58,-1.96,-1.65,1.65,1.96,2.58,1000)
```

```{r}
library(RColorBrewer)
MoranColours<- rev(brewer.pal(8, "RdGy"))
```

```{r}
# plot local Moran's I map
tm_shape(chicago_points_sf_joined) +
    tm_polygons("density_Iz",
        style="fixed",
        breaks=breaks1,
        palette=MoranColours,
        midpoint=NA,
        title="Local Moran's I, Density of Burglary")+
  tm_layout(legend.position = c(0.02,0.07),legend.title.size = 0.6, legend.text.size=.6)+
  tm_scale_bar(position = c(0.02, 0.005), size = 0.5,text.size = .75)+
  tm_compass(position = c(0.85, 0.8))
```

# descriptive analysis
```{r}
chicago_points_sf_joined$scale_density <- scale(chicago_points_sf_joined$density)
```
```{r}
chicago_points_sf_joined$lag_scale_density <- lag.listw(Lward.lw, chicago_points_sf_joined$scale_density)
```

```{r}
chicago_oas_moran_stats <- st_as_sf(chicago_points_sf_joined)
```

```{r}
# set a significance value
sig_level <- 0.1

# classification with significance value
chicago_oas_moran_stats$quad_sig <- ifelse(chicago_oas_moran_stats$scale_density > 0 & 
                                          chicago_oas_moran_stats$lag_scale_density > 0 & 
                                          I_LWard_Local_Density[,5] <= sig_level, 
                                          'high-high', 
                                   ifelse(chicago_oas_moran_stats$scale_density <= 0 & 
                                          chicago_oas_moran_stats$lag_scale_density <= 0 & 
                                          I_LWard_Local_Density[,5] <= sig_level, 
                                          'low-low', 
                                   ifelse(chicago_oas_moran_stats$scale_density > 0 & 
                                          chicago_oas_moran_stats$lag_scale_density <= 0 & 
                                          I_LWard_Local_Density[,5] <= sig_level, 
                                          'high-low', 
                                   ifelse(chicago_oas_moran_stats$scale_density <= 0 & 
                                          chicago_oas_moran_stats$lag_scale_density > 0 & 
                                          I_LWard_Local_Density[,5] <= sig_level, 
                                          'low-high',
                                   ifelse(I_LWard_Local_Density[,5] > sig_level, 
                                          'not-significant', 
                                          'not-significant')))))

# classification without significance value
chicago_oas_moran_stats$quad_non_sig <- ifelse(chicago_oas_moran_stats$scale_density > 0 & 
                                              chicago_oas_moran_stats$lag_scale_density > 0, 
                                              'high-high', 
                                       ifelse(chicago_oas_moran_stats$scale_density <= 0 & 
                                              chicago_oas_moran_stats$lag_scale_density <= 0, 
                                              'low-low', 
                                       ifelse(chicago_oas_moran_stats$scale_density > 0 & 
                                              chicago_oas_moran_stats$lag_scale_density <= 0, 
                                              'high-low', 
                                       ifelse(chicago_oas_moran_stats$scale_density <= 0 & 
                                              chicago_oas_moran_stats$lag_scale_density > 0,
                                              'low-high',NA))))
```

```{r}
# plot 1
# plot the results without the statistical significance
ggplot(chicago_oas_moran_stats, aes(x = scale_density, 
                                   y = lag_scale_density, 
                                   color = quad_non_sig)) +
  geom_vline(xintercept = 0) + # plot vertical line
  geom_hline(yintercept = 0) + # plot horizontal line
  xlab('Scaled Crime Density') +
  ylab('Lagged Scaled Crime Density') +
  labs(colour='Relative to neighbours') +
  geom_point()
```

```{r}
# plot 2
# plot the results with the statistical significance
ggplot(chicago_oas_moran_stats, aes(x = scale_density, 
                                   y = lag_scale_density, 
                                   color = quad_sig)) +
  geom_vline(xintercept = 0) + # plot vertical line
  geom_hline(yintercept = 0) + # plot horizontal line
  xlab('Scaled Crime Density') +
  ylab('Lagged Scaled Crime Density') +
  labs(colour='Relative to neighbours') +
  geom_point()
```

```{r}
# map all of the results here
tm_shape(chicago_oas_moran_stats) +
    tm_fill(col = 'quad_non_sig', palette = c("#de2d26", "#fee0d2", "#deebf7", "#3182bd"))+
   tm_layout(legend.position = c(0.02,0.08))+
  tm_scale_bar(position = c(0.02, 0.005), size = 0.5,text.size = .75)+
  tm_compass(position = c(0.85, 0.8))
```

```{r}
# map only the statistically significant results here
# map only the statistically significant results here
tm_shape(chicago_oas_moran_stats) +
    tm_fill(col = 'quad_sig', palette = c("#de2d26", "#fee0d2", "white")) +
    tm_borders(col = "grey")+
   tm_layout(legend.position = c(0.02,0.08))+
  tm_scale_bar(position = c(0.02, 0.005), size = 0.5,text.size = .75)+
  tm_compass(position = c(0.85, 0.8))
```


# use socoeconomic factors to do regression analysis
```{r}
density_data<-chicago_points_sf_joined%>%
  dplyr::select(area_num_1,density)%>%
  st_drop_geometry()
```

```{r}
density_data<-density_data%>%
  dplyr::mutate(density=round(as.numeric(density)*10000000+1,3))%>%
  dplyr::mutate(code=as.numeric(area_num_1))
```

```{r}
regression_data<-socioeconomic_factors%>%
  left_join(.,
            density_data,
            by=c("community_area_number"="code"))
```

# descpritive varibales
```{r}
regression_data%>%
dplyr::select(density)%>%
  pull()%>%
  qplot()+ 
  geom_histogram() 
```
```{r}
regression_data%>%
dplyr::select(percent_households_below_poverty)%>%
  pull()%>%
  qplot()+ 
  geom_histogram() 
```

```{r}
regression_data%>%
dplyr::select(per_capita_income)%>%
  pull()%>%
  qplot()+ 
  geom_histogram() 
```

```{r}
regression_data%>%
dplyr::select(percent_aged_16_unemployed)%>%
  pull()%>%
  qplot()+ 
  geom_histogram() 
```

```{r}
regression_data%>%
dplyr::select(percent_aged_25_without_high_school_diploma)%>%
  pull()%>%
  qplot()+ 
  geom_histogram() 
```

```{r}
regression_data%>%
dplyr::select(percent_of_housing_crowded)%>%
  pull()%>%
  qplot()+ 
  geom_histogram() 
```

```{r}
modelMLR <- lm(log(density) ~ log(percent_households_below_poverty)+log(percent_of_housing_crowded)+log(percent_aged_25_without_high_school_diploma)+log(percent_aged_16_unemployed)+log(per_capita_income), data = regression_data)
summary(modelMLR)
```

```{r}
chicago_points_sf_joined$residuals<-modelMLR$residuals
```


# residuals的map
```{r}
tm_shape(chicago_points_sf_joined) + tm_fill("residuals", style = "cont", palette = "-RdBu") +
tm_shape(chicago_shape) + tm_polygons(alpha = 0, border.alpha = 1, border.col = "black") +
tm_compass(position = c(0.85, 0.8)) +
tm_scale_bar(position = c(0.05, 0.005)) +
tm_layout(frame = FALSE, legend.title.size = 0.5, legend.text.size = 0.5)+
   tm_layout(legend.position = c(0.05,0.08))
```
# Global Moran's I for residuals
```{r}
#generate unique number for each row
chicago_points_sf_joined$ROWNUM <- 1:nrow(chicago_points_sf_joined)
# We need to coerce the sf spatialdatafile object into a new sp object
spatialdatafile_2.0 <- as(chicago_points_sf_joined, "Spatial")
# Create spatial weights matrix for areas
Weights <- poly2nb(spatialdatafile_2.0, row.names = spatialdatafile_2.0$ROWNUM)
WeightsMatrix <- nb2mat(Weights, style='B')
Residual_WeightMatrix <- mat2listw(WeightsMatrix , style='W')
# Run the test on the regression model output object "modelMLR" using lm.morantest()
lm.morantest(modelMLR, Residual_WeightMatrix, alternative="two.sided")
```

```{r}
descrptive_variables<-socioeconomic_factors[,c('per_capita_income','percent_households_below_poverty','percent_aged_25_without_high_school_diploma','percent_aged_16_unemployed','percent_of_housing_crowded')]
descrptive_variables2<-log(descrptive_variables)
```



```{r}
par(las = 1) # all axis labels horizontal

boxplot(as.data.frame(descrptive_variables2), main ="All Socialeconomic Factors",

      horizontal = TRUE)
```

```{r}
boxplot(as.data.frame(descrptive_variables2), main ="All Socialeconomic Factors")
```

```{r}
library(patchwork)
```

```{r}
p1<-ggplot(descrptive_variables2, 
                 aes(x=per_capita_income)) + 
  geom_histogram(color="black", 
                 fill="white")+
  labs(
       x="income", 
       y="Frequency")
p2<-ggplot(descrptive_variables2, 
                 aes(x=percent_households_below_poverty)) + 
  geom_histogram(color="black", 
                 fill="white")+
  labs( 
       x="households below poverty", 
       y="Frequency")
p3<-ggplot(descrptive_variables2, 
                 aes(x=percent_aged_25_without_high_school_diploma)) + 
  geom_histogram(color="black", 
                 fill="white")+
  labs(
       x="high school without diploma", 
       y="Frequency")
p4<-ggplot(descrptive_variables2, 
                 aes(x=percent_aged_16_unemployed)) + 
  geom_histogram(color="black", 
                 fill="white")+
  labs( 
       x="unemploy", 
       y="Frequency")
p5<-ggplot(descrptive_variables2, 
                 aes(x=percent_of_housing_crowded)) + 
  geom_histogram(color="black", 
                 fill="white")+
  labs( 
       x="housing crowded", 
       y="Frequency")
p1+p2+p3+p4+p5+plot_layout(ncol=3)
```